Typed Functional Genetic Programming

### 2013-07-16 tex/thesis.tex
[tex]

% !!! místo we can atd psát one can ...
% ujednotit i.e. , ty čaky kolem toho atd

@@

\section{Crossover}

@@

In GP the most common mutation is \textit{Subtree swapping mutation}.
This crossover randomly selects one node in each parent tree.
Two new child individuals are constructed by swapping subtrees 
which have roots in those selected nodes.\\

@@

\begin{todo}
 & říct že to je přesně ta notace co používá koza
 & EXAMPLES of tree representations of \lterms  
 & napsat transformace mezi \atTree a \sexprTree
 & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
   přímo \sexprTree a říct proč se tomu tak řiká 
   (na tom místě co je v textu dřív)

@@

\section{Crossover}

\begin{todo}
& Prostě budem křížit jen podstromy stejnýho typu
& ALE! Hlavní problém jsou proměnný
& ve stručnosti kouknout na dvě hlavní metody
  && křížit prostě lambda stromy a vzniklý problémy řešit online 
  && zbavit se proměnejch 
& říct že nakonec sme přilnuli k zbavit se proměnejch
\end{todo}

\subsection{Finding same types}
\subsection{Two basic options}
Resolve problems with free variables or avoid variables 
completely.

@@

@@

>>>>

% !!! místo we can atd psát one can ...
%      - ale ted to vypada že to neni tak žhavý pravidlo

% ujednotit i.e. , ty čaky kolem toho atd

@@

\section{Crossover}
\label{GPxover}

@@

In GP the most common crossover is \textit{Subtree swapping crossover}.
This crossover randomly selects one node in each parent tree.
Two new child individuals are constructed by swapping subtrees 
which have roots in those selected nodes.\\

@@

& ujistit se že je v části o tree reprezentacích 
  je napsaný co to znamená že podstrom má nějakej typ,
  že podstromy odpovídaj podvýrazum atd

 & říct že to je přesně ta notace co používá koza
 & EXAMPLES of tree representations of \lterms  
 & napsat transformace mezi \atTree a \sexprTree
 & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
   přímo \sexprTree a říct proč se tomu tak řiká 
   (na tom místě co je v textu dřív)

@@

\newpage
\section{Crossover}

The crossover operation in classical GP is performed 
by swapping randomly selected subtrees in each parent 
\sexprTree (see \ref{GPxover}).

For typed lambda terms two difficulties arise: Types and variables.

We will show how to crossover typed \lterm trees in both 
\atTree and \sexprTree notation.\\

As in classical GP our crossover will be performed by swapping
two subtrees. But now with constraint that both subtrees have
the same type. This is discussed in grater detail in \ref{typed-swapping}.\\

Variables bring more difficulties then types do.
This problem arises from variables that are free in subterms corresponding 
to swapped subtrees.

Following example illustrates the problem. \Lets have these two
parent trees with selected nodes in bold.\\

\newcommand{\lh}[1]{\lambda #1}

\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]

~\\The swap of subtrees results in following trees:\\

\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]

~\\The problem is that variable $x_2$ in second tree
is not bound by any $\lambda$-head and since
it is not element of $\Gamma$, the second tree is not well-typed \lterm.

% - This is only one example of a problem caused by swapping subtrees with variables,
% - Resolve problems with free variables or avoid variables completely.

In \ref{repairing-method} is presented more detailed discussion of problems with
variables and method which repairs such "broken" terms.

But we can avoid dealing with this problem by avoiding use of variables.
This can be achieved by process called abstraction elimination.

@@

\subsection{Typed subtree swapping}
\label{typed-swapping}

\subsection{ Repairing method }
\label{repairing-method}

\begin{todo}
& !!!
\end{todo}

\subsection{ Comparison of the two methods }
\label{comarison-ski-repairing}

\begin{todo}
& !!!
& říct že nakonec sme přilnuli k zbavit se proměnejch
\end{todo}

\newpage
\section{Mutation}

@@

\chapter{Polymorphic variant}

\begin{todo}
 & poznamky o čem se mimojiné zmínit:
   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
      jako alternativa k ADF ! Myslim že hodně zajimavý
\end{todo}
